% CREATED BY MAGNUS GUSTAVER, 2020
\thispagestyle{plain}			% Supress header
\section*{Acknowledgements}
I hereby want to express my deepest gratitude towards all the wonderful people who made my time at Chalmers and in Gothenburg, a truly unforgettable experience. 
Not only did all of them fill my journey with joy, but each person uniquely supported me in achieving this master's degree.
First, I want to deeply thank my examiner Tünde for giving me the opportunity to work on this thesis project at the Chalmers Plasma Theory group.
She creates an environment that enables everyone to flourish, both personally and academically. 
Tünde provided me with invaluable guidance for both my thesis and also for my scientific career in general.
Next, I am incredibly thankful towards my main supervisor, Oskar, who spent countless hours explaining and discussing in detail the ideas and concepts behind my thesis. 
Without Oskar, I would have never been able to carry out this project. 
Additionally, all of my supervisors, Oskar, István, Sarah and Tünde gave me fantastic feedback and advice during our weekly meetings and by thoroughly reviewing everything I have written in this thesis.
I could have never imagined receiving such excellent supervision, and I greatly appreciate all the time and effort they invested in supporting me.
I also want to thank Per Helander for allowing us to expand on his ideas about modelling the pellet rocket effect and for the fruitful discussions we had.
Overall, every member of the Chalmers Plasma Theory group made the everyday experience at the office very enjoyable, and I am grateful to each one of them.
Last but not least, I want to thank all of my friends and family, both here in Gothenburg and in Germany, for their support and for bringing immense amounts of happiness into my life. 

\vspace{1.5cm}
\hfill
\theauthor, Gothenburg, June 2024

\newpage				% Create empty back of side
\thispagestyle{empty}
\mbox{}